---
title: "GROUP ASSIGNMENTS 2021-04-28"
author: "NAMAZI | SCHMALHOFER | KNOCHE"
date: "04/28/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# loading packages
library(wooldridge)
# loading data sets 
data("wage2")
data("rdchem")
```

#### ASSIGNMENT C4
Use the data `wage2.dta` to estimate a simple regression explaining monthly salary (wage) in terms of IQ score (IQ)

1. Find the average salary and average IQ in the sample. What is the sample standard deviation of IQ? (IQ scores are standardized so that the average in the population is 100 with a standard deviation equal to 15.)   
```{r}
mean(wage2$wage) # average salary
mean(wage2$IQ)   # average IQ
sd(wage2$IQ)     # sample standard deviation of IQ
```

2. Estimate a simple regression model where a one-point increase in IQ changes wage by a constant dollar amount. Use this model to find the predicted increase in wage for an increase in IQ of 15 points. Does IQ explain most of the variation in wage?
```{r}
lm(wage ~ IQ, data = wage2)

wage = 116.992 + 8.303 * "IQ" 

wage = 116.992 + 8.303 * 15

wage = 8.303 * 15 

print(paste("delta-wage = delta-β0 + delta-β1 IQ + delta-u = 0 + 8.303 * 15 + 0 
            = 124.545"))
```

3. Now, estimate a model where each one-point increase in IQ has the same percentage effect on wage. If IQ increases by 15 points, what is the approximate percentage increase in predicted wage?
```{r}
log(wage) = β0 + β1 * IQ 
```

#### ASSIGNMENT C5
For the population of firms in the chemical industry, let rd denote annual expenditures on research and development, and let sales denote annual sales (both are in millions of dollars).   

1. Write down a model (not an estimated equation) that implies a constant elasticity between rd and sales. Which parameter is the elasticity?

2. Now, estimate the model using the data in `rdchem.raw`. Write out the estimated equation in the usual form. What is the estimated elasticity of rd with respect to sales? Explain in words what this elasticity means.