---
title: "GROUP ASSIGNMENTS 2021-05-05"
author: "NAMAZI | SCHMALHOFER | KNOCHE"
date: "05/05/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# loading packages
library(wooldridge)
# loading data sets 
data("bwght2")
```

## Chapter 3 - 5
In a study relating college grade point average to time spent in various activities, you distribute a survey to several students. The students are asked how many hours they spend each week in four activities: studying, sleeping, working, and leisure. Any activity is put into one of the four categories, so that for each student, the sum of hours in the four activities must be 168.

1. In the model $$GPA = \beta_{0} + \beta_{1} study + \beta_{2} sleep + \beta_{3} work + \beta_{4} leisure + u,$$ does it make sense to hold sleep, work, and leisure fixed, while changing study?

`No, it wouldn't make sense.`

2. Explain why this model violates Assumption MLR.3.

Because $GPA = \beta_{0} + \beta_{1} study + \beta_{2} sleep + \beta_{3} work + \beta_{4} leisure + u$, while is $GPA = 168$ means that $\beta_{1} study + \beta_{2} sleep + \beta_{3} work + \beta_{4} leisure \hat{=} 168$.
Accordingly any summand of $\beta_{1} study + \beta_{2} sleep + \beta_{3} work + \beta_{4} leisure$ is calculable through the equation.

3. How could you reformulate the model so that its parameters have a useful interpretation and it satisfies Assumption MLR.3?

`Leaving out any one coefficient (i.e. sleep) would suffice to satisfy MLR.3 .`

## Chapter 3 - C7
Use `bwght2.dta` to answer this question.

1. Estimate the model $$log(bwght) = \beta_{0} + \beta_{1} cigs + \beta_{2} npvis + u,$$
and report the results in the usual form, including the sample size and R-squared. Are the signs of the slope coefficients what you expected? Explain.

```{r echo=FALSE}
model_1 <- lm(lbwght ~ cigs + npvis, data = bwght2)
smry_1 <- model_1
smry_1
```

2. If `npvis` increases by one sample standard deviation, what is the estimated effect on birth weight?

```{r echo=FALSE}
sd_npvis = sd(bwght2$npvis, na.rm = TRUE) # cor( , use=)
sd_npvis
```

3. Now run the simple regression of log(`bwght`) on `cigs`, and compare the slope coefficient with the estimate obtained in part 1. Is the estimated effect of cigarette smoking much different than in part 1?


4. Find the correlation between `cigs` and `npvis` and use it to explain the similarity of the simple and multiple regression estimates of $\beta_{1}$.


5. Add the variables `mage`, `meduc`, `fage`, and `feduc` to the regression from part (i). Is birth weight [more precisely log(`bwght`)] an easy variable to explain?