---
title: "GROUP ASSIGNMENTS 2021-04-21"
author: "NAMAZI | SCHMALHOFER | KNOCHE"
date: "04/21/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# loading packages
library(wooldridge)
# loading data sets 
data("ceosal2")
data("sleep75")
```

#### ASSIGNMENT C2
The dataset `ceosal2.dta` contains information on chief executive officers for U.S. corporations. The variable `salary` is annual compensation, in thousands of dollars, and `ceoten` is prior number of years as company CEO.

1. Find the average salary and the average tenure in the sample.
```{r}
mean(ceosal2$salary)  # average salary
mean(ceosal2$ceoten)  # average tenure
```

2. How many CEOs are in the first year as CEO (that is, `ceoten = 0`)?.  
What is the longest tenure as a CEO?
```{r}
sum(ceosal2$ceoten == 0)  # CEOs in first year
max(ceosal2$ceoten)       # longest tenure as CEO
```

3. Estimate the simple regression model $log(salary) = \beta_{0} + \beta_{1}ceoten + u$.  
Report your results. What is the (approximate) predicted percentage increase in salary given one more year as a CEO?
```{r}
my_lm_salary <- lm(salary ~ ceoten, data = ceosal2)
my_lm_salary                  # regression model
my_lm_salary$coefficients[2]  # the predicted percentage increase in salary
```

#### ASSIGNMENT C3
Use the data `sleep75.dta` from Biddle and Hamermesh (1990) to study whether there is a tradeoff between the time spent sleeping per week and the time spent in paid work. We could use either variable as the dependent variable.  
For concreteness, estimate the model $sleep = \beta_{0} + \beta_{1}totwork + u$, where sleep is minutes spent spleeping at night per week and totwrk is total minutes worked during the week.

1. Report your results in equation form along with the number of observations and $R^{2}$.  
What does the intercept of the equation mean?
```{r}
my_lm_sleep <- lm(sleep ~ totwrk, data = sleep75)         # regression model
my_intercept_sleep <- my_lm_sleep$coefficients[1]
my_slope_sleep <- my_lm_sleep$coefficients[2]
my_summary_sleep <- summary(my_lm_sleep)
my_r_squared_sleep <- my_summary_sleep$r.squared          # R^2
```
report in equation
```{r echo=FALSE}
print(paste("sleep =", round(my_intercept_sleep, 4),
            "+ (", round(my_slope_sleep, 4), "*",
            names(my_slope_sleep), ") + u"))
```
form along with $R^{2}$
```{r echo=FALSE}
print(paste("R^2 =", round(my_r_squared_sleep, 4)))
```
The intercept of the equation corresponds to the amount of minutes per week that someone is expected to be sleeping at night, if they didn't work.

2. If totwrk increases by 2 hours, by how much is sleep estimated to fall?  
Do you find this to be a large effect?
```{r echo=FALSE}
print(paste("Sleep is estimated to change by 120 minutes *", names(my_slope_sleep),
            "=", round(120 * my_slope_sleep, 4), "minutes per week."))
```
In this case sleep would fall by 18.0895 minutes per week. That doesn't seem to be a large effect.